\section{Background: Object Algebras}\label{subsec:ObjectAlgebras}

Object Algebras capture a design pattern to solve the Expression
Problem~\cite{wadler98expression-problem}.  As a result Object
Algebras support modular and type-safe extensibility in two
dimensions: data variants and operations. Object Algebras can be used
conveniently to model recursive data types, such as ASTs. Here we
briefly introduce Object Algebras using the example of a simple
expression language.

\lstinputlisting[linerange=4-7]{../ObjectAlgebras/src/example_object_algebras/ObjectAlgebras.java} % APPLY:linerange=INT_ALG_INTERFACE

The above code shows an \emph{Object Algebra interface} which models
literal and addition expressions. The interface resembles an abstract
factory interface, but instead of constructing objects of a concrete
type, the result is abstract through the type parameter \lstinline{E}.
A concrete Object Algebra will implement such an interface by
instantiating the type parameter with a type that constructs objects
for a particular purpose. For instance, the class \lstinline{Eval}
(which implements evaluation of expression) is as follows:

\lstinputlisting[linerange=27-34]{../ObjectAlgebras/src/example_object_algebras/ObjectAlgebras.java} % APPLY:linerange=EXP_FACTORY

Using this algebra to create expressions will immediately evaluate
the expression to its result. To add another operation the interface
\lstinline{ExpAlg} is implemented again. As an example, the class
\lstinline{Print} presented below shows a simple printing operation on
expressions. Both examples illustrate extension in the dimension of
operations.

\lstinputlisting[linerange=38-43]{../ObjectAlgebras/src/example_object_algebras/ObjectAlgebras.java} % APPLY:linerange=INT_PRINT

%% \begin{figure}[t]
%% \nocaptionrule
%% \lstinputlisting[linerange=49-59]{../ObjectAlgebras/src/example_object_algebras/ObjectAlgebras.java} % APPLY:linerange=OA_TEST_CODE
%% \caption{Object Algebra Test Code}
%% \label{oa_test_code}
%% \end{figure}

The other dimension of extensibility (adding data variants) is
realized by extending the Object Algebra interface itself. For
instance, the interface \lstinline{MulAlg} defined below extends the
original \lstinline{ExpAlg} with multiplication expressions.

\lstinputlisting[linerange=11-13]{../ObjectAlgebras/src/example_object_algebras/ObjectAlgebras.java} % APPLY:linerange=INT_BOOL_ALG_INTERFACE

\noindent The key feature of Object Algebras is that the existing operations
(evaluation and printing) can now be extended to support
multiplication without having to change the existing \lstinline{Eval}
and \lstinline{Print} classes. For example, evaluation is extended 
wuth multiplication as follows:

\lstinputlisting[linerange=17-22]{../ObjectAlgebras/src/example_object_algebras/ObjectAlgebras.java} % APPLY:linerange=MUL_EVAL

In summary, Object Algebra interfaces describe recursive data types,
and implementations of those interfaces represent
operations. Extending the interface allows developers to extend the
data type with new data variants, and (re)implementing the interface
allows developers to define operations.  Finally, extension in both
dimensions is fully type-safe and does not compromise separate
compilation. Although for space reasons this section does not show
how to use Object Algebras, such examples will be provided throughout
the paper.

%Examples that show how to use Object Algebras are provided 


%What makes \emph{object algebras} a useful design pattern even more is that it does not rely on any advanced language features, and it is both extensible and type-safe.


\begin{comment}

This paper showed how various types of traversals for complex
structures can be automatically provided by \Name. \name traversals are
written directly in Java and are type-safe, extensible and separately
compilable. There has always been a tension between the
correctness guarantees of static typing, and the flexibility of
untyped/dynamically-typed approaches. \name shows that even
in type systems like Java's, it is possible to get considerable
flexibility and adaptability for the problem of boilerplate code in
traversals of complex structures, without giving up modular static typing.
%\name offers the best of both worlds: the
%static typing guarantees; and flexibility and adaptability.

There are many of avenues for future work. One area of research is to
extend \name traversals to support flexible traversal strategies,
similarly to strategic
programming~\cite{borovansky1996elan,visser1998core,vandenBrand:2003:TRT:941566.941568}. Another
line of work worth exploring is to adopt generalizations of object
algebras~\cite{oliveira13fop} for added expressiveness of \name
traversals.

\end{comment}
