\documentclass[10pt,nocopyrightspace,preprint]{sigplanconf}

\usepackage{color}

\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{pstricks,pst-node,pst-tree}
\usepackage{theorem}
\usepackage{bbm}
\usepackage{pgf}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{xspace}
\usepackage{ql}
\usepackage{mdwlist}
\usepackage{morefloats}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
\usepackage{dirtree}
\usepackage{booktabs}

\lstset{ %
language=Java,                % choose the language of the code
columns=flexible,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numbers=none,                   % where to put the line-numbers
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
deletekeywords={label}
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}

\pagestyle{plain}


\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand\jason[1]{\authornote{jason}{green}{#1}}
\newcommand\haoyuan[1]{\authornote{haoyuan}{red}{#1}}
\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
\newcommand\tijs[1]{\authornote{tijs}{blue}{#1}}

\newcommand\name{{\bf Shy}\xspace}
\newcommand\Name{{\bf Shy}\xspace}

\newcommand{\hl}[1]{\textcolor{red}{#1}}

\newcommand\sem[1]{\llbracket #1 \rrbracket_r}
\newcommand\sems[1]{\llbracket #1 \rrbracket_s}
\newcommand\tsem[1]{\llbracket #1 \rrbracket}
\newcommand{\rbm}[1]{\raisebox{-2.0ex}[0.5ex]{#1}}
\newcommand\nat[0]{\mathbb{N}}
\newcommand\unit[0]{\mathbbm{1}}

\begin{comment}
\author{Author1$^{\textrm{1}}$ Author2$^{\textrm{2}}$ }
\institute{$^{\textrm{1}}$The University of Hong Kong\\
\email{author1@cs.hku.hk}\\
$^{\textrm{2}}$ The University of Hong Kong\\
\email{author2@cs.hku.hk}}
\authorrunning{Author1 and Author2} % abbreviated author list (for running head)
\end{comment}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\begin{comment}
\conferenceinfo{SPLASH '15}{October 25--30, 2015, Pittsburgh, Pennsylvania, USA.}
\copyrightyear{2015}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish,
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers,
                                  % short abstracts)

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.
\end{comment}

\title{Scrap your Boilerplate with Object Algebras}
%\subtitle{Subtitle Text, if any}

\authorinfo{Name1}
           {Affiliation1}
           {Email1}
\authorinfo{Name2\and Name3}
           {Affiliation2/3}
           {Email2/3}

\maketitle

\begin{abstract}

Traversing complex \emph{Abstract Syntax Trees} (ASTs) typically requires
large amounts of tedious boilerplate code. For many operations most of
the code simply walks the structure, and only a small portion of the
code implements the functionality that motivated the traversal in the
first place.  This paper presents a type-safe Java framework called
\name that removes much of this boilerplate code. In \name
\emph{Object Algebras} are used to describe complex and extensible
AST structures. Using Java annotations \name generates generic boilerplate code
for various types of traversals. For a concrete traversal, users of
\name can then inherit from the generated code
and override only the interesting cases. Consequently, the amount of code that users
need to write is significantly smaller. Moreover, traversals using the
\name framework are also much more \emph{structure shy}, becoming more
adaptive to future changes or extensions to the AST structure.
%An additional benefit of \name is that traversal code can be easily
%reused by extensions of the
To prove the effectiveness of the approach, we applied \name
in the implementation of a domain-specific questionnaire
language. Our results show that for a large number of traversals
there was a significant reduction in the amount of user-defined code.
%\bruno{Say something more about extensibility and type-safety!}

\begin{comment}

Static types are useful to distinguish between different kinds
of nodes in a structure and to prevent misuses. The
distinction between different types of nodes also means that code for
dealing with each type of node is needed. However, in large structures,
such as the Abstract Syntax Tree (AST) of a programming language,
the amount of required code can be a problem. For some operations,
which traverse large structures, most of the code amounts to recursively
delegating the traversal to the child nodes. Only for some nodes
of the structure the code needs to do something different. Still
the programmer needs to diligently and tediously write the error-prone
traversal code for all nodes.

In this paper we present a framework

This approach works well for many operations,
which need different (and non-trivial) code for each different type
of node

%Unfortunately, for some operations the interesting

%%In large structures,
%%such as the Abstract Syntax Tree (AST) of a programming language,
%%the code required to traverse the whole structure is proportionaly large.


The problem is particularly prominent
in statically typed languages, where the typing discipline
enforces strict distinctions between different cases

Operations that traverse complex structures often require large and
tedious amounts of boilerplate code. In those operations there are
typically a few

A pervasive problem in programming occurs when large tree tr

\end{comment}

\end{abstract}

%\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore,
% you may leave them out
%\terms
%term1, term2

%\keywords
%keyword1, keyword2

%if False

%\category{D.3.2}{Programming Languages}
%                {Language Classifications}
%                [Functional Languages]
%\category{F.3.3}{Logics and Meanings of Programs}
%                {Studies of Program Constructs}
%                []

%\terms
%Languages
%
%\keywords
%Mixins, explicit effects, monads, aspect-oriented programming, parametricity,
%interference

%endif

%===============================================================================

\input{sections/Introduction.tex}

\input{sections/ObjectAlgebras.tex}

\input{sections/Overview.tex}

\input{sections/Queries.tex}

\input{sections/GeneralizedQueries.tex}

\input{sections/Transformations.tex}

\input{sections/ContexualTransformations.tex}

\input{sections/Extensible.tex}

%%\input{sections/Framework.tex}

\input{sections/CaseStudy.tex}

\input{sections/RelatedWork.tex}

\input{sections/Conclusion.tex}


\bibliographystyle{abbrvnat}
\bibliography{paper}


\appendix
\input{sections/Appendix.tex}
\begin{comment}
\section{Appendix Title}

This is the text of the appendix, if you need one.

\acks

Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

\begin{thebibliography}{}
\softraggedright

\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
P. Q. Smith, and X. Y. Jones. ...reference text...

\end{thebibliography}
\end{comment}

\end{document}
