\section{Introduction}

Many applications require complex recursive data structures. Examples
abound, for example, in language processing tools/libraries for programming
languages, domain-specific languages, mark-up languages like HTML, or
data-interchange languages like XML or JSON. In those applications
Abstract Syntax Trees (ASTs) are the key data structure needed to
model the various constructs of these languages. Such ASTs have various
different types of nodes, which can range from a few dozen to
several hundred kinds of nodes (for example in the ASTs of languages like Java).
\tijs{This is probably not true... Should we mention Cobol?}

Static types are helpful to deal with such complex structures.  Static types formalize the distinction between different kinds of
nodes. Furthermore the distinctions are helpful to ensure that
traversals over these structures have an appropriate piece of code that
deals with each different type of node. This can prevent a large class
of run-time errors that would not otherwise be detected.

%For example static types prevent errors that would arize from
%applying the code intended for a certain kind of node to another kind
%of node. Static types can also ensure that all nodes are dealt with by
%an appropriate piece of code.

Unfortunately, when traversing such structures, the number of nodes and the
enforced type distinctions between nodes can lead to so-called
\emph{boilerplate code}~\cite{ralf03syb}: code that is similar for most types of nodes and which
essentially just walks the structure. Traversals where such boilerplate
code dominates are called \emph{structure shy}~\cite{DemeterBook}. In
structure shy operations the ``interesting'' code is limited to only a small portion of nodes.
A typical example is computing the free
variables of an expression in some programming language. In this
case, the interesting code occurs in the nodes representing the
binding and variable constructs. In all other cases, the code would just deal with
walking the structure. In data structures with dozens or hundreds of
kinds of nodes, having to explicitly write code for each kind of node is both tedious and error-prone.
%In summary, structure shy operations
%suffer from the static typing discipline, because the type
%distinctions between different kinds of nodes leads to significant
%boilerplate code.

The boilerplate problem in implementing traversals has
received considerable attention in the past. For example, both
\emph{Adaptive Object-Oriented Programming} (AOOP)~\cite{DemeterBook}
and \emph{Strategic
  Programming}~\cite{borovansky1996elan,visser1998core}
are aimed partly at solving this problem. Most approaches to AOOP and
strategic programming use some meta-programming techniques, such as
code generation or reflection. The use of meta-programming offers
programmers an easy way to avoid having to write boilerplate code.
This has important benefits. Firstly the user has
to write much less code, also removing the possibility of errors in
the code walking the structure. Secondly the code becomes much more
adaptive to changes: if a change to the traversed data type  only affects the generated boilerplate code,  the user-defined code
can remain unchanged.  However such
meta-programming based approaches  usually come at the cost of
other desirable properties, such as type-safety, extensibility or
separate compilation. The functional programming community has also
studied the problem before. For instance,  the popular
``Scrap your boilerplate''~\cite{ralf03syb} approach supports type-safety and separate
compilation.  However most of the techniques used in functional
languages cannot be easily ported to mainstream OO languages like Java, and are
limited in terms of extensibility.

This paper presents a Java framework called \name that allows users to
define \emph{type-safe and extensible structure-shy operations}. \name uses
\emph{object algebras}~\cite{bruno12oa} to describe complex data structures. Object
algebras are a recently introduced technique, which has been shown to
have significant advantages for software extensibility.  In \name
object algebra interfaces are combined with Java annotations to
generate generic and reusable object algebras that deal with
boilerplate traversal code.  Those object algebras include different
types of traversals: \emph{queries}; \emph{transformations};
\emph{generalized queries}; and \emph{contextual transformations}. Programmers
who want to implement structure-shy traversals can inherit the
generic traversal code, and focus only on writing the interesting
parts of the traversal. Consequently, the amount of code that
programmers need to write is significantly smaller.
Traversals
written in \name are:

\begin{itemize}

\item {\bf Adaptive and structure shy:} \name traversals can omit
  boilerplate code, allowing these traversals to be more adaptive to
  future changes or extensions to the data type.

\item {\bf Simple and general:} \name traversals work for any
  structure that can be expressed as a (multi-sorted) object
  algebra. This includes complex OO hierarchies or ASTs for large
  languages. Very often traversals are quite simple, being
  implementable in just a few lines of code, even for complex
  structures with hundreds of different types of nodes.

\item {\bf Implemented in plain Java:} \name traversals do not require
  a new tool or language. The approach is library based and uses only
  Java annotations.

\item {\bf Type-safe with separate compilation:} \name traversals are directly written in Java
  and the Java type-system ensures type-safety. No run-time casts are
  needed for generic traversal code or for user-defined traversal
  code. Furthermore \name traversals support separate compilation.

\item {\bf Extensible:} Traversals inherit type-safe
  extensibility from object algebras. Both traversals and structures
  are extensible. Therefore it is possible to
  reuse traversal code in structures that are extended with additional
  node types.

%\item {\bf Performant:} Traversals written with \name have performance that
%is comparable to naive hand-written OO code.

\end{itemize}

To prove the effectiveness of the approach, we have applied \name
in the implementation of the domain-specific questionnaire
language QL~\cite{gouseti14extensible}.
Our results show that for a large number of traversals
there was a significant reduction in the amount of user-defined code.

Although we have chosen Java as the implementation language for \Name,
our approach should apply to any OO language with support for generics
and annotations. For example it should be easy to port \name to
languages such as Scala or C\#.

\begin{comment}

Existing work:

dynamic approaches: they can offer adaptivity and
structure shyness, but lack type-safety.

static approaches: No fully type-safe approach that we know
of for Java. Some approaches that rely on reflection and/or casts.
But these can have run-time type-errors and performance can be
quite bad.
Some static type-safe in the functional programming community,
but cannot be easily ported to Java (rely on sophisticated type
system features; target algebraic data types; ...). Moreover
(with a few exceptions) most approaches do not offer extensibility.

Static types are useful to distinguish between different kinds
of nodes in a structure and to prevent misuses. The
distinction between different types of nodes also means that code for
dealing with each type of node is needed. However, in large structures,
such as the Abstract Syntax Tree (AST) of a programming language,
the amount of required code can be a problem. For some operations,
which traverse large structures, most of the code amounts to recursively
delegating the traversal to the child nodes. Only for some nodes
of the structure the code needs to do something different. Still
the programmer needs to diligently and tediously write the error-prone
traversal code for all nodes.

In this paper we present a framework

This approach works well for many operations,
which need different (and non-trivial) code for each different type
of node

%Unfortunately, for some operations the interesting

%%In large structures,
%%such as the Abstract Syntax Tree (AST) of a programming language,
%%the code required to traverse the whole structure is proportionaly large.


The problem is particularly prominent
in statically typed languages, where the typing discipline
enforces strict distinctions between different cases

Operations that traverse complex structures often require large and
tedious amounts of boilerplate code. In those operations there are
typically a few

A pervasive problem in programming occurs when large tree tr

\end{comment}

In summary, the contributions of this paper are:

\begin{itemize}

\item {\bf Design patterns for generic traversals.} We provide a set of design
patterns for various types of traversals using object algebras. These include:
\emph{queries}, \emph{transformations},
\emph{generalized queries} and \emph{contextual transformations}.

\item {\bf The Shy Java framework.} We have implemented\footnote{{\bf
      Note to Reviewers:} Due to the anonymous review process we
    submit a file bundle with the implementation together with the paper submission. If the
    paper is accepted we will make the code publicly available.} a
  Java framework that can be used to describe complex structures using
  object algebras; and to eliminate boilerplate code. The framework
  uses Java annotations to automatically generate generic traversals.


\item {\bf Case study and empirical evaluation.} We evaluate the
  approach using a case study based on the QL domain-specific
  language. The results of our case study show significant savings in
  terms of user-defined traversal code.
\end{itemize}
