\documentclass[preprint]{llncs}
\usepackage{etex}

%\usepackage{color}

%\usepackage{amsmath}
%\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url}
%\usepackage{bbm}
%\usepackage{pgf}
%\usepackage{multirow}
\usepackage{listings}
\usepackage{verbatim}
%\usepackage{graphicx}
%\usepackage{wrapfig}
\usepackage{xspace}
\usepackage{ql}

\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
%\usepackage{dirtree}

\lstset{ %
language=Java,                % choose the language of the code
columns=flexible,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numbers=none,                   % where to put the line-numbers
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
morekeywords={var},
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}

\pagestyle{plain}


\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand\jason[1]{\authornote{jason}{green}{#1}}
\newcommand\haoyuan[1]{\authornote{haoyuan}{red}{#1}}
\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
\newcommand\tijs[1]{\authornote{tijs}{blue}{#1}}

\newcommand\name{{\bf Shy}\xspace}
\newcommand\Name{{\bf Shy}\xspace}

\newcommand{\hl}[1]{\textcolor{red}{#1}}

\newcommand\sem[1]{\llbracket #1 \rrbracket_r}
\newcommand\sems[1]{\llbracket #1 \rrbracket_s}
\newcommand\tsem[1]{\llbracket #1 \rrbracket}
\newcommand{\rbm}[1]{\raisebox{-2.0ex}[0.5ex]{#1}}
\newcommand\nat[0]{\mathbb{N}}
\newcommand\unit[0]{\mathbbm{1}}

\author{Hoayuan Zhang$^{\textrm{1}}$ Jason Chu$^{\textrm{2}}$ }
\institute{$^{\textrm{1}}$The University of Hong Kong\\
\email{author1@cs.hku.hk}\\
$^{\textrm{2}}$ The University of Hong Kong\\
\email{author2@cs.hku.hk}}
\authorrunning{Haoyuan Zhang et al.} % abbreviated author list (for running head)


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%\conferenceinfo{POPL'10} {}
%%\CopyrightYear{2010}


\title{Scrap your Boilerplate with Object Algebras}

\maketitle

\begin{abstract}

Traversing complex data structures typically requires large amounts of
tedious boilerplate code. For many operations most of the code simply
walks the structure, and only a small portion of the code implements
the functionality that motivated the traversal in the first place.
This paper present a type-safe Java framework called \name that
removes much of this boilerplate code. In \name \emph{object algebras}
are used to describe complex \emph{extensible} data structures. Using
Java annotations generic boilerplate code is generated for various
types of traversals, including queries and transformations.  Thus,
using \Name, programmers can inherit the generic traversal code to
focus only on writing the interesting parts of the
traversals. Consequently, the amount of code that programmers need to
write is significantly smaller. Moreover, traversals using the \name framework
are also much more \emph{structure shy}, becoming more adaptive
to future changes or extensions to the data structure.
%An additional benefit of \name is that traversal code can be easily
%reused by extensions of the
To prove the effectiveness of the approach, we employeed \name
on the implementation of a domain-specific questionaire
language. Our results show that for a large number of traversals
there was a significant reduction in the amount of user-defined code.
%\bruno{Say something more about extensibility and type-safety!}

\begin{comment}

Static types are useful to distinguish between different kinds
of nodes in a structure and to prevent misuses. The
distinction between different types of nodes also means that code for
dealing with each type of node is needed. However, in large structures,
such as the Abstract Syntax Tree (AST) of a programming language,
the amount of required code can be a problem. For some operations,
which traverse large structures, most of the code amounts to recursively
delegating the traversal to the child nodes. Only for some nodes
of the structure the code needs to do something different. Still
the programmer needs to diligently and tediously write the error-prone
traversal code for all nodes.

In this paper we present a framework

This approach works well for many operations,
which need different (and non-trivial) code for each different type
of node

%Unfortunately, for some operations the interesting

%%In large structures,
%%such as the Abstract Syntax Tree (AST) of a programming language,
%%the code required to traverse the whole structure is proportionaly large.


The problem is particularly prominent
in statically typed languages, where the typing discipline
enforces strict distinctions between different cases

Operations that traverse complex structures often require large and
tedious amounts of boilerplate code. In those operations there are
typically a few

A pervasive problem in programming occurs when large tree tr

\end{comment}

\end{abstract}

%===============================================================================

\input{sections/Introduction.tex}

\input{sections/Overview.tex}

\input{sections/Queries.tex}

\input{sections/GeneralizedQueries.tex}

\input{sections/Transformations.tex}

\input{sections/ContexualTransformations.tex}

\input{sections/Extensible.tex}

\input{sections/Framework.tex}

\input{sections/CaseStudy.tex}

\input{sections/RelatedWork.tex}

\input{sections/Conclusion.tex}


\bibliographystyle{splncs}
\bibliography{paper}

\appendix

\end{document}
